---
title: "R Notebook"
output: html_notebook
---

#Libraries
```{r}
library(tidyverse)
```

#Data
```{r}
books <- read_csv("books.csv")
```

#Data exploration
Find out what kind of information is in the dataset and what questions are 
appropriate to ask.


```{r}
#Explore the dataset
str(books)
dim(books)
names(books)
#Find out if there are any NAs in the dataset
sum(is.na(books))
#Find out whether there are outliers in the numeric variables
summary(books)
skimr::skim(books)

```
The _num_pages_, _ratings_count_ and _text_reviews_count_ variables have 
outliers where the values are very high, meaning I should not use the mean
value when working with these variables as it has clearly been effected by 
high outliers.

Language is not likely to be of much use since Goodreads is more likely to be
used by readers in English-speaking countries, so I can't make any conclusions
about language anyway as it will be biased toward English books.

##Actions
-Explore zero values in the _num_pages_, _ratings_count_ and _average_ratings_
variables
-Calculate outliers - outliers for _average_ratings_ could tell us which books
are extremely highly rated and extremely lowly rated. 
-Remove books with zero ratings since we're interested in ratings


#Exploring missing values
There aren't any missing values, but there are some variables with a value
of zero

```{r}
#Explore books with zero pages and those with less than 10 pages b/c we would 
#expect even children's books to have more than 10 pages
books %>% 
  select(title, num_pages, average_rating, ratings_count) %>% 
  filter(num_pages == 0)

books %>% 
  select(title, num_pages, average_rating, ratings_count) %>% 
  filter(num_pages < 10)
```

Some books with zero pages are audiobooks, some are collections of books and
some are unclear why they would be listed as having zero pages. There
doesn't seem to be a connection between books with under 10 pages and the 
number of ratings and average rating.

```{r}
#Explore books with 0 ratings and then a low number of ratings
books %>% 
  select(title, num_pages, average_rating, ratings_count) %>% 
  filter(ratings_count == 0)
```

Some books with zero ratings have an average rating above zero, which doesn't
make any sense. This enforces the case that all books with 0 ratings should
be removed from the dataset.

```{r}
#Explore books with a zero average rating and a rating of 5
books %>% 
  select(title, num_pages, average_rating, ratings_count) %>% 
  filter(average_rating == 0)

books %>% 
  select(title, num_pages, average_rating, ratings_count) %>% 
  filter(average_rating == 5)
```
All books with a zero average rating also don't have any ratings, so they can
be removed without risking removing books that everyone just hated.

All books with a rating of 5 have a very low number of ratings (n <= 5), so 
they should also be removed because ratings with a very low number of raters
can't be trusted. We should consider the pros and cons of removing all books
with fewer than 5 raters.

I looked up the formula for finding outliers in average rating to find very low
rated books and very high rated books. This number is also useful to find very
low rated books with very few raters and very high rated books with very few 
raters. I can then exclude these from the dataset.
The formula for finding low outliers is Quartile 1 - 1.5(Inter-quartile range)
The formula for finding high outliers is Quartile 3 + 1.5 (Interquartile range)
Low outliers < 3.215
High outliers > 4.695

```{r}
#Explore books with between 1 and 10 raters and avg ratings which are outliers
books %>% 
  select(title, num_pages, average_rating, ratings_count) %>% 
  filter(between(ratings_count, "1", "10")) %>% 
  filter(average_rating < 3.215 | average_rating > 4.695)
```

I will lose 82 books if I exclude these from the dataset, but with a dataset
with over 10000 observations, this won't matter very much and will also exclude
potentially unreliable data, so I think the benefits outweigh any potential 
negatives. I want to investigate the highest and lowest rated books, so it 
would be beneficial to exclude very high and very low rated books which also 
have a very low number of raters.

##Actions
-Select variables title, average_ratings, ratings_count, authors
-Filter out books with zero ratings
-Filter out books with outlier ratings that have fewer than 10 raters.

#Questions to ask
-Which books are the highest and lowest rated? Group these into categories.
-Which Tolkien book is rated the highest?
-Find books that are about Tolkien - e.g. contain  Tolkien, "Lord of the Rings",
"Middle Earth", "Hobbit" in the title but not Tolkien as the author

#Select variables and clean the data

```{r}
#Select variables that relate to questions
book_ratings <- books %>% 
  select(title, authors, average_rating, ratings_count)
book_ratings
```

```{r}
#Exclude books with zero ratings
book_ratings <- book_ratings %>% 
  filter(ratings_count != 0)
```

```{r}
#Filter out books with 0 ratings
book_ratings %>% 
  filter(ratings_count == 0)
```

```{r}
#Filters out books that have fewer than 10 raters and an average rating that is
#a high or low outlier
book_ratings <- book_ratings %>% 
   filter(!((ratings_count < 10) & 
              (average_rating < 3.215 | average_rating > 4.695)))
```

#Question 1: Which books have the highest and lowest average ratings? Group
them in categories and find out how many books are in each category.
##Highest rated: Existential Meditation, The Complete Calvin and Hobbes,
The Feynman Lectures on Physics Vols 7-8, Harry Potter Boxed Set Books 1-5,
It's a Magical World (Calvin and Hobbes #11).
##Lowest rated: Citizen Girl, The Trouble with Pears, A Matter of Trust,
The Governess; or The Little Female Academy, Alentejo Blue

```{r}
book_ratings %>% 
  slice_max(average_rating, n = 20)

book_ratings %>% 
  slice_min(average_rating, n = 20)

summary(book_ratings) #Finds the summary stats for the new dataset
```
##Rating Categories
The categories I'll be grouping the books into are Very High, High, Medium, Low,
and Very Low. Very high and very low are the high and low outliers. High are
above the 3rd quartile value, medium are between the 3rd and 1st quartile 
values and Low are below the 1st quartile value.


```{r}
book_ratings <- book_ratings %>% 
  mutate(
    average_ratings_cat = case_when(
      average_rating > 4.695 ~ "Very high",
      average_rating > 4.140 ~ "High",
      average_rating > 3.780 ~ "Medium",
      average_rating >= 3.215 ~ "Low",
      average_rating < 3.215 ~ "Very Low"
    )
  ) %>% 
  arrange(desc(average_rating))
```

```{r}
#Counts the number of books in each rating category
book_ratings %>% 
  count(average_ratings_cat)
```
```{r}
#Finds the median number of ratings for each category
book_ratings %>% 
  group_by(average_ratings_cat) %>% 
  summarise(avg_rating_num = median(ratings_count))
```



#Question 2: Which Lord of the Rings book is rated the highest? 
The boxed sets are rated the highest, but the highest rated book in the 
trilogy is Return of the King

```{r}
#Filter all books that contain Tolkien in the author column
book_ratings %>% 
  filter(grepl("Tolkien", authors, ignore.case = TRUE))
```

#Question 3: -Find books that are about Tolkien but not by him - e.g. contain  
Tolkien, "Lord of the Rings", "Middle Earth", "Hobbit" in the title but not 
Tolkien as the author

```{r}
book_ratings %>% 
  filter((grepl("Lord of the Rings", title, ignore.case = TRUE)) |
         (grepl("Middle-earth", title, ignore.case = TRUE)) |
           (grepl("Hobbit", title, ignore.case = TRUE)) |
           (grepl("Tolkien", title, ignore.case = TRUE))) %>% 
  filter(!(grepl("Tolkien", authors, ignore.case = TRUE)))
```

# Notes from homework recap

```{r}
#Find the authors with more than 20 books and rank them by their average rating
books %>%
  #splits strings that have two or more authors separated by "/"
  mutate(authors = str_split(authors, "/")) %>% 
  #Duplicates info for each separate author(?)
  unnest(authors) %>% 
  group_by(authors) %>%
  summarise(author_count = n(),
         author_average_rating = mean(average_rating)) %>%
  filter(author_count > 20) %>%
  arrange(desc(author_average_rating))
```

```{r}
books %>% 
  distinct(publisher) %>% 
  arrange(publisher)
```

