---
title: "R Notebook"
output: html_notebook
---
##Libraries
```{r}
library(tidyverse)
```

##Read in data
```{r}
movie_data <- read_csv("movies.csv")
movie_data
```
##Question 1
The first thing I noticed is that the data is formatted correctly (no metadata),
so we don't need to skip some rows. It looks like there are two variables for 
title so I will need to choose between the two. The other variables are named 
sensibly, so I don't need to change their names. There is some info that will 
probably not be very useful for summary stats such as the summary of the movie 
and the movie's tagline. We haven't learned any text analysis yet, so I 
probably can't do anything with these. We should check for NAs and also for 
unexpected values (e.g. where runtime is 0 or where the budget is 0).

```{r}
summary(movie_data)
```
##Question 2 - Select variables of interest

```{r}
movies_selected <- movie_data %>% 
  select(title, runtime, budget)
```

##Question 3 - Find missing values

```{r}
summary(is.na(movies_selected))
```

##Question 4 - Convert "0" runtime into NAs in new variable "movies_runtime" 
and then count NAs again

```{r}
movies_selected <- movies_selected %>% 
  mutate(movies_runtime = na_if(runtime, "0"),
         .after = runtime)
  summary(is.na(movies_selected))
```


##Question 5 - Impute missing values in runtime with median runtime and check
for missing values

```{r}
movies_imputed <- movies_selected %>% 
  mutate(runtime = coalesce(movies_runtime, 
                            median(
                              movies_runtime, na.rm = TRUE)))
movies_imputed
```
```{r}
summary(is.na(movies_imputed))
```

##Question 6 - Find movies with 10 shortest and 10 longest runtimes using 
_slice_min_ and _slice_max_

```{r}
movies_imputed %>% 
  slice_min(runtime, n = 10)

movies_imputed %>% 
  slice_max(runtime, n= 10)
```

##Question 7 - Use the _if_else_ function to replace budgets below $100 with
the median budget

```{r}
movies_imputed <- movies_imputed %>% 
  mutate(budget = if_else(
    budget < 100, 
    median(budget, na.rm = TRUE), 
    budget))
movies_imputed
```

##Check that there are no budgets below 100 in new dataset
```{r}
summary(movies_imputed)
```

##Extention 1 - Use _case_when_ to create a new budget type column

```{r}
movies_imputed <- movies_imputed %>% 
  mutate(
    movie_budgets = case_when(
      budget < 12000000                   ~ "Small Budget",
      between(budget, 12000000, 40000000) ~ "Medium Budget",
      budget > 40000000                   ~ "Big Budget"
    )
  )
```

##Check count for each group in category movie_budgets
```{r}
movies_imputed %>% 
  count(movie_budgets)
```

