---
title: "R Notebook"
output: html_notebook
---

# Libraries

```{r}
library(tidyverse)
library(GGally)
library(ggfortify)
library(fastDummies)
```

# 1) Load the housing_prices.csv data set and undertake an initial exploration of the data. 

```{r}
housing_prices <- read_csv("data/housing_prices.csv")
```

```{r}
housing_prices %>% head()
housing_prices %>% summary()
```

# 2) We expect the total_rooms of houses to be strongly correlated with total_bedrooms. Use ggpairs() to investigate correlations between these two variables.

```{r}
ggpairs(housing_prices, columns = 4:5)
```

The correlation between the two variables is very strong at 0.934.

# 3) So, we do find significant correlations. Let’s drop total_bedrooms from the dataset, and use only total_rooms going forward.

```{r}
housing_prices <- housing_prices %>% 
  select(-total_bedrooms)
```

# 4) We are interested in developing a regression model for the median_house_value of a house in terms of the possible predictor variables in the dataset.

## 4.1) Use ggpairs() to investigate correlations between median_house_value and the predictors (this may take a while to run, don’t worry, make coffee or something).

```{r}
ggpairs(housing_prices)
```

## 4.2) Perform further ggplot visualisations of any significant correlations you find.

```{r}
#latitude and median house value
housing_prices %>% 
  ggplot(aes(x = latitude,
             y = median_house_value)) +
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)

```

```{r}
#longitude and median house value
housing_prices %>% 
  ggplot(aes(x = longitude,
             y = median_house_value)) +
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
#housing_median_age and median house value
housing_prices %>% 
  ggplot(aes(x = housing_median_age,
             y = median_house_value)) +
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
#total rooms and median house value
housing_prices %>% 
  ggplot(aes(x = total_rooms,
             y = median_house_value)) +
  geom_point()+
  scale_x_sqrt()+
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
#households and median house value
housing_prices %>% 
  ggplot(aes(x = households,
             y = median_house_value)) +
  geom_point()+
  scale_x_sqrt() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
#median_income and median house value
housing_prices %>% 
  ggplot(aes(x = median_income,
             y = median_house_value)) +
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)
```

# 5) Shortly we may try a regression model to fit the categorical predictor ocean_proximity. Investigate the level of ocean_proximity predictors. How many dummy variables do you expect to get from it?

I expect to get 4 dummy variables (n - 1)

```{r}
housing_prices %>% 
  group_by(ocean_proximity) %>% 
  summarise(number = n())
```

# 6) Start with simple linear regression. Regress median_house_value on median_income and check the regression diagnostics.

```{r}
model_income_value <- lm(median_house_value ~ median_income, housing_prices)

autoplot(model_income_value)

summary(model_income_value)
```
# 7) Add another predictor of your choice. Check your assumptions, diagnostics, and interpret the model.

```{r}
#add dummies for ocean_proximity and clean them up
housing_dummies <- housing_prices %>% 
  fastDummies::dummy_cols(
    select_columns = "ocean_proximity",
    remove_first_dummy = TRUE,
    remove_selected_columns = TRUE
  ) %>% janitor::clean_names() %>% 
  mutate(across(ocean_proximity_inland : ocean_proximity_near_ocean, as.logical))
```

```{r}
#visualise near_bay on box plot and look at correlation
housing_dummies %>% 
  ggplot(aes(ocean_proximity_near_bay, median_house_value))+
  geom_boxplot()

housing_dummies %>% 
  summarise(cor = cor(ocean_proximity_near_bay, median_house_value))
```

```{r}
#model using multiple linear regression
model_income_bay <- lm(median_house_value ~ 
                         median_income + ocean_proximity_near_bay,
                       housing_dummies)

#check diagnostics
autoplot(model_income_bay)# remember to ask about this, looks like there is a downward trend in the first plot.

summary(model_income_bay)
```

For every 10K increase in median income, we can expect to see median house value increase by $39,611.70 on average (holding all other variables constant).

If the house block is located near the bay, we can expect the median house value to increase by $40,158.20 on average (holding all other variables constant)

R^2 is 0.429, so 42.9% of the variation in median_house_value is explained by proximity to a bay and median income.