---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
```

```{r}
library(CodeClanData)
```
Question 1 - Load the dataset and look at it. What are the dimensions, 
variable types, variable names?
```{r}
drinks_content
```
#Dimensions of dataset - 242 rows and 18 columns
```{r}
dim(drinks_content)
```
#Variable types and names - a mix of numeric and categorical variables 
containing types of beverages at Starbucks and their nutritional information.
```{r}
glimpse(drinks_content)
```
#Question 2 - Select variables and check for NA values - no NA values in calories
```{r}
coffee_calories <- select(drinks_content, Beverage_category, Beverage, 
                          Beverage_prep, Calories)
```
```{r}
coffee_calories
```
```{r}
sum(is.na(coffee_calories$Calories))
```
#Question 3 - create new variable calorie_diff which takes away 135 calories from each drink

```{r}
coffee_calories %>% 
mutate(calorie_diff = Calories - 135)
```

#Question 4 - Summarise the mean number of calories in each beverage category
Smoothies, Frappucino Blended Coffee and Signature Espresso have 
the most calories on average
Coffee, Shaken Iced Beverages and Classic Espresso drinks have the least
calories on average

#Create a group_by variable called Beverage_category_grouped
```{r}
Beverage_category_grouped <- group_by(coffee_calories, Beverage_category)
```

#Summarise the mean number of calories for Beverage_category_grouped and arrange
in ascending order.
```{r}
summarise(Beverage_category_grouped, Avg_calories = mean(Calories)) %>% 
  arrange(Avg_calories)
```



#Question 5 - What is the average number of calories in each Beverage_prep type?

#Group dataset by Beverage_prep type, summarise the average number of calories
for each group, arrange by number of calories.
```{r}
coffee_calories %>% 
  group_by(Beverage_prep) %>% 
  summarise(Average_Calories = mean(Calories)) %>% 
  arrange(Average_Calories)
```

#Question 6 - Which Beverage_prep type contains more than the average calories 
of all drinks?
5 beverage prep types contain more than the average calories of all drinks
1) Whole Milk
2) Venti Nonfat Milk
3) 2% Milk
4) Grande Nonfat Milk
5) Soymilk

#First, find the avg calories of all drinks and then convert into numeric value
using pull

```{r}
average_calories_all_drinks <- coffee_calories %>% 
  summarise(average_value = mean(Calories)) %>% 
  pull()
average_calories_all_drinks
```
#Then create a summary table grouped by beverage prep with an extra column 
that finds the difference between average calories for the beverage_prep type
and the average calories in all the drinks. Filter out any values that are
negative so that we only see which beverage_prep types contain more than
the average calories for all drinks.

```{r}
coffee_calories %>% 
  group_by(Beverage_prep) %>% 
  summarise(Average_Calories = mean(Calories)) %>% 
  mutate(avg_calorie_diff = Average_Calories - average_calories_all_drinks) %>% 
  filter(avg_calorie_diff > 0)
```

#Question 7 - Which is the best type of coffee drink to get if you're worried
about consuming too many calories?
A short brewed coffee b/c it only has 3 calories

Filter data to only include drinks with the Beverage_category "Coffee" and 
arrange by Calories in ascending order
```{r}
coffee_calories %>% 
  filter(Beverage_category == "Coffee") %>% 
  arrange(Calories)
```

