---
title: "R Notebook"
output: html_notebook
---
# Libraries

```{r}
library(tidyverse)
library(CodeClanData)
library(scales)
```

# Data

```{r}
qb_revenue <- qb_revenue_breakdown
qb_monthly_sales <- qb_monthly_sales
qb_competitors <- qb_competitors
qb_device <- qb_device_data
```



# 2.1 Part 1

## Q1: Take the data in the dataset qb_revenue_breakdown and make a stacked bar chart showing the sources of revenue across the two years in the dataset.

```{r}
qb_revenue %>% 
  ggplot(aes(x = Year,
             y = Revenue,
             fill = Product)) + 
  geom_col()
  labs(x = "Year",
       y = "Revenue",
       title = "Sources of revenue in 2018 and 2019")
```

## Q2: Make a line chart showing monthly sales of the “flagship product” - the ‘KwikBit Unit 1’ - over the last year. This data is inside qb_monthly_sales, but you will need to use filter() or subset() (base R).

```{r}
qb_monthly_sales %>% 
  filter(Cashflow == "Kwikbit Unit 1 Sales") %>% 
  ggplot(aes(x = Date,
             y = Amount)) +
  geom_line() 
  labs(x = "Date",
       y = "Amount",
       title = "Monthly Sales of Kwikbit Unit 1",
       subtitle = "Data from monthly sales\n")
```

## Q3: Make a line chart showing monthly revenue and costs over the last year. This data is also in qb_monthly_sales.

```{r}
qb_monthly_sales %>% 
  filter(Cashflow != "Kwikbit Unit 1 Sales") %>% 
  ggplot(aes(x = Date,
             y = Amount,
             colour = Cashflow)) +
  geom_line() +
  labs(x = "Date",
       y = "Amount",
       title = "Monthly revenue and costs for 2018/2019",
       fill = "Cash flow")
```

## Q4: Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use geom_area). This data is in qb_competitors.

```{r}
qb_competitors %>% 
  ggplot(aes(x = Year,
             y = Revenue,
             fill = Company)) +
  geom_area() +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Competitor fitness tracker sales for the past 5 years")
```

## Q5: Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.

```{r}
qb_competitors %>% 
  ggplot(aes(x = Year,
             y = Revenue,
             colour = Company)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Competitor fitness tracker sales for the past 5 years")
```

## Q6: Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in qb_device_data.

```{r}
qb_device %>% 
  group_by(device, id, hours) %>% 
  summarise(total_steps_counted = sum(counts)) %>% 
  ggplot(aes(x = hours,
         y = total_steps_counted)) +
  geom_line() +
  facet_grid(device ~ id) +
  labs(x = "\nTime tracked",
       y = "Steps counted",
       title = "Number of steps counted over time on different devices",
       subtitle = "Comparison between Qikbit and top competitors for 5 participants\n")
```

# 2.2 Part 2

## Q1 - Take the plots that you produced in part one and now polish them by:

1. Adding appropriate labels
Edited above

2. Changing the scales and coordinates when appropriate.
3. Applying a unified theme, which is described below:

Graphs should have white backgrounds, and use colour sparingly.
There should be faint grid lines.
Font sizes should be ~12pt although titles should be slightly larger and axis labels can be slightly smaller.
All plots should use colours from the following company colour scheme.


```{r}

qb_theme <- function() {
  theme(
    #add white background
    panel.background = element_blank,
    #modify grid
    panel.grid.major.x = element_line(colour = "#B6A7B5",
                                      size = 0.25),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(colour = "#B6A7B5",
                                      size = 0.5),
    panel.grid.minor.y = element_blank(),
    #axis text slightly smaller than default
    axis.text = element_text(size = rel(0.75)),
    #axis titles slightly larger than default
    axis.title = element_text(size = rel(1.5)),
    
    
    
    
  )
}
```

