---
title: "R Notebook"
output: html_notebook
---

## Libraries

```{r}
library(gtools)
library(tidyverse)
library(janitor)
source("../../../Codeclan_work/Week_06/Probability/prob.R")
```
## Data


```{r}
titanic_data <- read_csv("titanic.csv")
```
## 1.1 - The Titanic


### Question
Calculate the conditional probabilities prob(survived∣crew) and prob(survived∣passenger) to settle this dispute.

Have a look at the distinct values of the class variable, this will tell you whether each individual was a passenger (1st, 2nd or 3rd class) or a member of the crew (treat restaurant staff as crew)



```{r}
names(titanic_data)

titanic_data %>% 
  distinct(class)

#make a new variable that designates passengers/crew
titanic_data <- titanic_data %>% 
  mutate(passenger_crew = if_else(str_detect(class, "[:digit:]"),
                                  "Passenger",
                                  "Crew"))
```

prob(survived|passenger) = 0.3796507 or 37.96%
prob(survived|crew) = 0.2370787 or 23.7%

Passengers were more likely to survive than crew

```{r}
#get contingency table with percentages to check calculations below.
titanic_data %>% 
  tabyl(passenger_crew, survived) %>% 
  adorn_percentages(denominator = "all") %>% 
  adorn_totals(where = c("row", "col"))


titanic_data %>% 
  #filter to only include passengers
  filter(passenger_crew == "Passenger") %>% 
  #divide the sum of passengers who survived by the total number of passengers.
  summarise(prob = sum(survived == "yes") / n())

titanic_data %>% 
  #filter to only include crew
  filter(passenger_crew == "Crew") %>% 
  #divide the sum of crew who survived by the total number of crew
  summarise(prob = sum(survived == "yes") / n())
```




## 1.2 Probability consolidation

### Q1 - If I roll a six-sided dice once, what is the probability of getting a number less than 5

4/6 or 2/3

### Q2 - How many outcomes are there in the sample space of an experiment in which we toss three dice

216

```{r}

rolldie(3)
```

### Q3 - In an experiment where we roll four dice:

    - how many outcomes are in the event where the total of the four dice 
    equals 5, 6 or 7)?: 34
    - what is the probability of this event?: 0.02623457
    - what assumptions do we apply in answering part ii? That the die is 6-sided,
      that all dice are fair dice (e.g. there is a 1/6 of rolling any number 1:6),
      that we can repeat numbers and it doesn't matter what order they are rolled
      in.

```{r}
roll_die_four <- rolldie(4) 

roll_die_four %>% 
  mutate(die_sum = rowSums(.)) %>% 
  filter(die_sum %in% 5:7) %>% 
  summarise(prob = n() / nrow(roll_die_four))
```

### Q4 - Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

There are 120 different ways they could sit in the 5 chairs and there is only 
one way that fits the event above. So the probability is: 0.008333333

```{r}
permutations(n = 5, r = 5, repeats.allowed = FALSE) %>% 
  nrow()

1/120
```

### Q5 - I roll six dice:

what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right?

1 in 46656 or 0.00002143347

what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order?
720/46656 = 0.0154321

```{r}
permutations(n = 6, r = 6, repeats.allowed = TRUE) %>% 
  nrow()
options(scipen = 99)
1/46656

roll_die_six <- rolldie(6)
permutations(n = 6, r = 6, repeats.allowed = FALSE) %>% 
  nrow()
720/46656
```

